---
title: "Introduction to Geospatial Techniques for Social Scientists in R"
subtitle: "Introduction to GIS"
author: "Stefan Jünger & Dennis Abel"
date: April 09, 2025
institute: |
  GESIS Workshop
presenter: Stefan
--- 
layout: true

```{r child = "./content/config/sessions_setup.Rmd"}
```

---

## About This Course

This course will teach you how to exploit `R` and apply its geospatial techniques in a social science context.

By the end of this course, you should... 

- Be comfortable with using geospatial data in `R`
  - Including importing, wrangling, and exploring geospatial data
- Be able to create maps based on your very own processed geospatial data in `R`
- Feel prepared for (your first steps in) spatial analysis
 
---

## Keep Calm and Carry On Learning/Using `R`

```{r, learn-r-cartoon, out.width = "50%", echo = F}
woRkshoptools::include_picture("r_first_then.png") 
```
<small><small>Illustration by [Allison Horst](https://github.com/allisonhorst/stats-illustrations) </small></small> 

---

## We are (Necessarily) Selective

There's a multitude of spatial `R` packages
- We cannot cover all of them
  - and we cannot cover all functions
- You may have used some we are not familiar with

We will show the use of packages we exploit in practice
- There's always another way of doing things in `R`
- Don't hesitate to bring up your solutions

**You can't learn everything at once, but you also don't have to!**

---

## Prerequisites for This Course

.pull-left[
- Knowledge of `R`, its syntax, and internal logic
  - Affinity for using script-based languages
  - Don't be scared to wrangle data with complex structures
- Working versions of `R` (and `Rstudio`) on your computer 
  - Ideally, with the packages installed, we asked you upfront
]

.pull-right[
<br>
<br>
```{r gates, echo = FALSE, out.width="75%"}
woRkshoptools::include_picture("i-love-programming.jpg")
```
]
  
---

## About Us

**Stefan Jünger**

.pull-left[
```{r me-pic, echo = FALSE, out.width = "50%"}
woRkshoptools::include_picture("Stefan.png")
```
]

.pull-right[
.small[
- Senior Researcher in the team Survey Data Augmentation at the GESIS department Survey Data Curation
- Ph.D. in social sciences, University of Cologne
]
]

.small[
- Research interests:
  - Quantitative methods, Geographic Information Systems (GIS)
  - Social inequalities
  - Attitudes towards minorities
  - Environmental attitudes
  - Reproducible research
]

.small[
[stefan.juenger@gesis.org](mailto:stefan.juenger@gesis.org) |
[https://stefanjuenger.github.io](https://stefanjuenger.github.io)
]

---

## About Us

**Anne Stroppe**

.pull-left[
```{r Anne-pic, echo = FALSE, out.width = "50%"}
woRkshoptools::include_picture("Anne.png")
```
]

.pull-right[
- Research Associate in the SDA team at GESIS
- Pursuing Ph.D. at the University of Mannheim (PolSci)
]


- Research focuses:
  - Political and electoral geography
  - Attitudes towards the political system
  - Political engagement and conspiracy beliefs
  - Data quality and open science
   
.small[
[anne-kathrin.stroppe@gesis.org](mailto:anne-kathrin.stroppe@gesis.org) 
]

---

## About You

.pull-left[
- What's your name?  
- Where do you work/research? What are you working on/researching?
- What is your experience with `R` or other programming languages?
- Do you already have experience with geospatial data?
]

.pull-right[
<br>
```{r tell-me, echo = FALSE}
woRkshoptools::include_picture("so-tell-me-5be02d.jpg")
```
.tinyisher[.center[[Source](https://makeameme.org/meme/so-tell-me-5be02d)]]
]
- **What do you want to use `R` and geospatial data for?**

---

## Preliminaries

- The workshop consists of a combination of lectures and hands-on exercises
- Feel free to ask questions at any time
- Slides and other materials are available at

.center[`https://github.com/StefanJuenger/gesis-workshop-geospatial-techniques-R-2024`]

---

## Download Materials

```{r download-repo, echo = FALSE, out.width = "75%"}
woRkshoptools::include_picture("download_repo.png")
```

---

## Course Schedule

```{r course-content, echo = FALSE}
course_content
```

---

## Now

```{r course-content-now, echo = FALSE}
course_content %>%
  kableExtra::row_spec(1, background = "yellow")
```

---

## Geographic Information in Social Science Research

.pull-left[
Exploiting geographic information is not new.

For example, Siegfried (1913) used soil composition information to explain election results in France.
]

.pull-right[
```{r siegfried-pic, echo = FALSE, out.width = "75%"}
woRkshoptools::include_picture("siegfried.png")
```
.center[.tinyisher[https://images.fr.shopping.rakuten.com/photo/874882994.jpg]]
]

---

## The Chicago School!

```{r park-etal-pic, echo = FALSE, out.width = "50%"}
woRkshoptools::include_picture("park_etal.png")
```
.tinyisher[Park et al. 1925]

---

## Integration in Theory

.pull-left[
```{r allport-pic, echo = FALSE, out.width = "65%"} 
woRkshoptools::include_picture("nature_prejudice.png")
```
.center[.tinyisher[https://i.gr-assets.com/images/S/compressed.photo.goodreads.com/books/1387595670l/7034925.jpg]]
]

.pull-right[
A lot of classic theories inherently make use of space.

For example, in Allport's (1954) contact theory, he at least implicitly assumes that people meet in a set space.
]

---

## Still relevant today

So many studies still rely on these ideas but incorperate space directly, e.g.,

.pull-left[
.tinyisher[
- Iyer, A., & Pryce, G. (2023). Theorising the causal impacts of social frontiers: The social and psychological implications of discontinuities in the geography of residential mix. Urban Studies, 00420980231194834. https://doi.org/10.1177/00420980231194834
- Kent, J. (2022). Can urban fabric encourage tolerance? Evidence that the structure of cities influences attitudes toward migrants in Europe. Cities, 121, 103494. https://doi.org/10.1016/j.cities.2021.103494
- Schmidt, K., Jacobsen, J., & Iglauer, T. (2023). Proximity to refugee accommodations does not affect locals’ attitudes toward refugees: Evidence from Germany. European Sociological Review, jcad028. https://doi.org/10.1093/esr/jcad028
- Xu, A. Z. (2023). Segregation and the Spatial Externalities of Inequality: A Theory of Interdependence and Public Goods in Cities. American Political Science Review, 1–18. https://doi.org/10.1017/S0003055423000722
- ...
- Jünger, S., & Schaeffer, M. (2023). Ethnic Diversity and Social Integration—What are the Consequences of Ethnic Residential Boundaries and Halos for Social Integration in Germany? KZfSS Kölner Zeitschrift Für Soziologie Und Sozialpsychologie. https://doi.org/10.1007/s11577-023-00888-1
]
]

.pull-right[
```{r pic-xu, echo = FALSE, out.width = "60%"}
woRkshoptools::include_picture("/xu2023.png")
```
.center[.tinyisher[Xu 2023]]
]

---

## Data Landscape

.pull-left[
Increased amount of available data
- Quantitative and on a small spatial scale

</br>

```{r pic-tools, echo = FALSE, out.width = "60%"}
woRkshoptools::include_picture("/tools_today.png")
```
]

.pull-right[
```{r pic-geo, echo = FALSE}
woRkshoptools::include_picture("geo_today.png")
```

Better tools
- Personal computer with enough horsepower
- Standard software, such as `R`, can be used as Geographic Information System (GIS)
]

---

## What Are Geospatial Data?

.pull-left[
</br>
Data with a direct spatial reference $\rightarrow$ **geo-coordinates**
- Information about geometries
- Optional: Content in relation to the geometries
]

.pull-right[
```{r pic-geometries-1, echo = FALSE, out.width = "85%"}
woRkshoptools::include_picture("fig_geometries.png")
```

.tinyisher[Sources: OpenStreetMap / GEOFABRIK (2018), City of Cologne (2014), and the Statistical Offices of the Federation and the Länder (2016) / Jünger, 2019]
]

---

## Geospatial Data in This Course I

In the folder called `./data`, you can find (most of) the data files prepped for all the exercises and slides. The following data are included:

- Administrative borders of Germany (Prefix *VG250_*) are provided by the German [Federal Agency for Cartography and Geodesy](http://www.bkg.bund.de)(2018). Check out their [Open Data Portal](https://gdz.bkg.bund.de/index.php/default/open-data.html).

- The locations of e-car charging stations in Germany are retrieved from the [Bundesnetzagentur](https://www.bundesnetzagentur.de/DE/Fachthemen/ElektrizitaetundGas/E-Mobilitaet/Ladesaeulenkarte/start.html). 

- Election Results for the German Right-Wing Populist Party *AfD* in the 2021 German federal election ([Der Bundeswahlleiter, Wiesbaden 2018](https://www.bundeswahlleiter.de/bundestagswahlen/2017/ergebnisse/weitere-ergebnisse.html)).

---

## Geospatial Data in This Course II

- District attributes are from [INKAR](https://www.inkar.de/) and curated by Bundesinstitut für Bau-, Stadt- und Raumforschung (BBSR)

- German Census 2011 data are provided by the [Federal Statistical Office Germany, Wiesbaden 2020](https://www.zensus2011.de/EN/Home/home_node.html)

- Shapefiles, voting data, and car-related data for Cologne are gathered from the [Open Data Portal Cologne](https://www.offenedaten-koeln.de/)

- Information about public transport stops and streets are from [OpenStreetMap](https://www.openstreetmap.org/)

- Weather data are gathered from [Climate Data Center of the German Weather Service](https://www.dwd.de/DE/klimaumwelt/cdc/cdc_node.html)

**Please make sure that if you reuse any of the provided data to cite the original data sources.**

---

## What Is GIS?

Most common understanding: Geographic Information Systems (GIS) as specific software to process geospatial data for 
- Visualization
- Analysis

```{r pic-gis, echo = FALSE, out.width = "70%", fig.align = "center"}
woRkshoptools::include_picture("gis.png")
```

.center[.tinyisher[Screenshot of the Open Source GIS [`QGIS`](https://qgis.org)]]

---

## Data Specifics

.pull-left[
</br>
</br>
```{r pic-geometries-2, echo = FALSE}
woRkshoptools::include_picture("fig_3d_simple.png")
```

.tinyisher[Sources: OpenStreetMap / GEOFABRIK (2018) and City of Cologne (2014)]
]

.pull-right[
Formats
- Vector data (points, lines, polygons)
- Raster data (grids)

Coordinate reference systems
- Allow the projection on earth's surface
- Differ in precision for specific purposes
]

---

## Layers Must Match!

.pull-left[
.center[
(EPSG:3857)
]
]

.pull-right[
.center[
(EPSG:3035)
]
]

```{r pic-crs, echo = FALSE}
woRkshoptools::include_picture("fig_projections.png")
```

.tinyisher[Source: Statistical Office of the European Union Eurostat (2018) / Jünger, 2019]

---

## Types of CRS

Differentiating between different CRS is wild (at least for me...). You may hear from geographic, geocentric, projected, or local CRS in your research.

What's the difference?

- whether 2 dimensional (longitude, latitude) or 3 dimensional (+height) coordinates are used
- the location of the coordinate system's origin (center of earth or not)
- projection on a flat surface (transformation of longitudes and latitudes to x and y coordinates)
- location (the smaller, the more precise are projections)

*In practice, you shouldn't worry too much about CRS. Again, what matters is that they match.*

---

class: middle

<br>

```{r layers-meme, echo = FALSE, out.width="70%"}
woRkshoptools::include_picture("layers_meme.jpg")
```

---

## Old Standard: `PROJ.4` Strings

This is how your information about the CRS are defined in a classic standard:

```
+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs 
```

.tinyisher[Source: https://epsg.io/3035]

(It's nothing you would type by hand)

---

## New Kid in Town: `WKT` (“Well Known Text”)

```
PROJCS["ETRS89 / LAEA Europe",
    GEOGCS["ETRS89",
        DATUM["European_Terrestrial_Reference_System_1989",
            SPHEROID["GRS 1980",6378137,298.257222101,
                AUTHORITY["EPSG","7019"]],
            TOWGS84[0,0,0,0,0,0,0],
            AUTHORITY["EPSG","6258"]],
        PRIMEM["Greenwich",0,
            AUTHORITY["EPSG","8901"]],
        UNIT["degree",0.0174532925199433,
            AUTHORITY["EPSG","9122"]],
        AUTHORITY["EPSG","4258"]],
    PROJECTION["Lambert_Azimuthal_Equal_Area"],
    PARAMETER["latitude_of_center",52],
    PARAMETER["longitude_of_center",10],
    PARAMETER["false_easting",4321000],
    PARAMETER["false_northing",3210000],
    UNIT["metre",1,
        AUTHORITY["EPSG","9001"]],
    AUTHORITY["EPSG","3035"]]
```

.tinyisher[Source: https://epsg.io/3035]

---

## Current Transition From `PROJ.4` to `WKT` in `R`

There's a transition from the old standard of `PROJ.4` to `WKT` in many `R` packages.
- sometimes, you'll get warnings
- you should check what's causing the warning
- in the best case, they will disappear in the future

---

## EPSG Codes

.pull-left[
Eventually, it's not as challenging to work with CRS in `R` as it may seem
- we don't have to use PROJ.4 or WKT strings directly

Most of the times it's enough to use so-called EPSG Codes ("European Petroleum Survey Group Geodesy")
- Small digit sequence
]

.pull-right[
<br>
```{r epsg-icon, echo = FALSE}
woRkshoptools::include_picture("ImageEPSGlogo.gif")
```
.tinyisher[.center[[Source](http://geotiff.maptools.org/proj_list/guid7.html)]]
]

---

## More Details on Geospatial Data

Let's learn about geospatial data as we learn about specific formats

```{r geo-real-worls, echo = FALSE, out.width="50%"}
woRkshoptools::include_picture("Raster-and-Vector-data-types-as-representative-of-real-world-geographic-space_W640.jpg")
```

.tinyisher[.center[[Source](https://dx.doi.org/10.13140/2.1.3030.1767)]]

---

## Packages in This Course

We will use plenty of different packages during the course, but only a few are our main drivers (e.g., the `sf` package). Here's the list of packages

.pull-left[
- [`dplyr`](https://cran.r-project.org/web/packages/dplyr/index.html)
- [`ggplot2`](https://cran.r-project.org/web/packages/ggplot2/index.html)
- **[`ggsn`](http://oswaldosantos.github.io/ggsn/)**
- [`haven`](https://cran.r-project.org/web/packages/haven/index.html)
- [`osmdata`](https://cran.r-project.org/web/packages/osmdata/index.html)
- **[`OpenStreetMap`](https://cran.r-project.org/web/packages/OpenStreetMap/index.html)**\*
- [`sf`](https://cran.r-project.org/web/packages/sf/index.html)
- [`spatialreg`](https://cran.r-project.org/web/packages/spatialreg/index.html)
]

.pull-right[
- [`spatstat`](https://cran.r-project.org/web/packages/spatstat/index.html)
- [`sfdep`](https://cran.r-project.org/web/packages/sfdep/index.html)
- [`spdep`](https://cran.r-project.org/web/packages/spdep/index.html)
- [`terra`](https://cran.r-project.org/web/packages/terra/index.html)
- **[`tidygeocoder`](https://cloud.r-project.org/web/packages/tidygeocoder/index.html)**
- [`tmap`](https://cran.r-project.org/web/packages/tmap/index.html)
- [`tmaptools`](https://cran.r-project.org/web/packages/tmaptools/index.html)
- [`raster`](https://cran.r-project.org/web/packages/raster/index.html)
- [`z11`](https://github.com/StefanJuenger/z11)
]

.tinyisher[\* If you use a Mac, you may need to install Java: https://www.java.com/de/download/]

---

class: middle
## Exercise 1_1_1: Package Installation

[Exercise](https://stefanjuenger.github.io/gesis-workshop-geospatial-techniques-R-2024/exercises/1_1_1_Package_Installation.html)

[Solution](https://stefanjuenger.github.io/gesis-workshop-geospatial-techniques-R-2024/solutions/1_1_1_Package_Installation.html)


<!-- STOP WITH SLIDES HERE -->

---

```{r child = "./content/config/sessions_end.Rmd"}
```